---
title: "Explore Joining"
output: html_notebook
---


```{r}
pacman::p_load(rio, tidyverse)
beneficiary <- import("data/beneficiary_summary.rds")
rx <- import("data/prescription_drug_events.rds")
inpatient <- import("data/inpatient_claims.rds")
outpatient <- import("data/outpatient_claims.rds")

reference_sex <- tibble(cd = c(1, 2), desc = c("Male", "Female"))
```

Task 01

**Goal**: Compare the number of inpatient stays in year 2008 of men to women. 

To do this, we will break this down into two big steps.

- This first step does not require any relational logic.
- Create a new data set called `inpatient_counts`.
    - Use `inpatient`.
    - Filter to `clm_yr` equal to 2008.
    - Group by `desynpuf_id`, which is our beneficiary identifier.
    - Count the number of rows to get the number of unique patients.
- Remember, it is easier to write this one piece at a time and confirm that it works than to write everything all at once and discover you made multiple errors.
- Watch out of red x's. If you see one, it means there is an error on that line or possibly the line above.

```{r}
inpatient_counts <-
  inpatient |>
  filter(clm_year == 2008) |>
  group_by(desynpuf_id) |>
  count()
inpatient_counts
```

Record these values as you go through this document to make it easier for yuo to complete Canvas.

- **Question:** How many rows are in inpatient_counts now?
- **Answer:** 

Once you are satisfied you have the right answer, use the `bene_sex_ident_cd` from beneficiaries to add a new column to our data set.


- I recommend starting with `beneficiaries` and left joining to `inpatient_counts`.
- The left join function is actually smart enough to recognize the common column name, so you don't have to write that in unless the message annoys you.
- But you will have to figure out how to filter beneficiaries data to 2008 first.
- Extra credit if you handle the patients with ZERO inpatient utilization.
- Once you figure out the join, overwrite inpatient_counts with this new data set. 
- When you are done, you should have the following columns in `inpatient_counts`:
  - `desynpuf_id`,
  - `bene_sex_ident_cd`,
  - `n`
  - You can have others if you want, but you need these.

```{r}
inpatient_counts <-
  beneficiary |>
  filter(year == 2008) |>
    left_join(inpatient_counts) |>
  mutate(
    bene_sex_ident_cd = as.factor(bene_sex_ident_cd),
    n = replace_na(n, 0)
  ) |>
  select(desynpuf_id, bene_sex_ident_cd, n)
```

- **Question:** How many rows are in inpatient_counts now?
- **Answer:** 


# Task 02
- Now, let's look at the distribution.
- Draw density plot of n, the number of inpatient utilizations.

```{r}
ggplot(data = inpatient_counts, aes(x = n)) +
  geom_density()
```

- Now apply a facet wrap so that you can compare men and women.

```{r}
ggplot(data = inpatient_counts, aes(x = n)) +
  geom_density() +
  facet_wrap(~bene_sex_ident_cd)
```



# Task 03

- What is the minimum number of stays?
- What is the maximum number of stays?

```{r}
inpatient_counts |>
  summarize(
    min_n = min(n),
    max_n = max(n)
  )
```

# Task 04
- Insurance providers often try to provide preventative services to their most expensive patients.
- For example, if we look at the top 1% of patients utilizing inpatient services:
  - This code will give us the cut-off.
```{r}
inpatient_counts |>
  summarize(cut_point = quantile(n, probs = .99))
```

So if we know the cut-off is 3, then we can flag this utilization.

```{r}
inpatient_counts <-
  inpatient_counts |>
  mutate(
    ## hu is short for high utilizers
    hu = if_else(n > 3, TRUE, FALSE)
  )
```

This means that patients with more than 3 inpatient stays account for roughly 1% of the beneficiaries.

But what percentage of inpatient stays do these patients account for?

- Above we created a column called hu.
- Can you calculate the percentage of inpatient stays caused by these high utilizers?
```{r}
inpatient_counts |>
  group_by(hu) |>
  summarize(n = n()) |>
  ungroup() |>
  mutate(p = 100 * n / sum(n))

inpatient_counts |>
  group_by(hu) |>
  summarize(total_utilization = sum(n)) |>
  ungroup() |>
  mutate(p = 100 * total_utilization / sum(total_utilization))
```

You should see that the top 1% of patients cause much more than 1% of inpatient stays.


Calculate the percentage of high utilizers who are male?

```{r}
inpatient_counts |>
  filter(hu == TRUE) |>
  group_by(bene_sex_ident_cd) |>
  count() |>
  ungroup() |>
  mutate(p = 100 * n / sum(n))
```

```{r}
inpatient_counts |>
  group_by(bene_sex_ident_cd) |>
  count() |>
  ungroup() |>
  mutate(p = 100 * n / sum(n))
```

```{r}
reference_sex <-
  reference_sex %>%
  mutate(cd = as.factor(cd))

inpatient_counts %>%
  filter( hu == TRUE) %>%
  left_join(reference_sex, by = c("bene_sex_ident_cd" = "cd")) %>%
  group_by(desc) %>%
  count() %>%
  ungroup() %>%
  mutate(percentage = 100 * n / sum(n))



reference_sex <- 
  reference_sex |>
  mutate(cd = as.factor(cd))

inpatient_counts |>
  filter(hu == TRUE) |>
  left_join(
    reference_sex,
    by = c("bene_sex_ident_cd" = "cd")
  ) |>
  group_by(desc) |>
  count() |>
  ungroup() |>
  mutate(p = 100 * n / sum(n))
```


