---
title: "Explore Polypharmacy"
output: html_notebook
---


```{r}
# This is the only code I'm going to give you for this.
pacman::p_load(rio, tidyverse)
beneficiary <- import("data/beneficiary_summary.rds")
rx <- import("data/prescription_drug_events.rds")
```

# Joining

## 00000B48BCF4AD29

- Normally, we treat ids as PHI and NOT include it in at report like this.
- But this is fine. 🤷
- Join by desnypuf_id and clm_year
  - clm_year is a variable I added during import.

### Beneficiary Data

```{r}
beneficiary |>
  filter(desynpuf_id == "00000B48BCF4AD29")
```

### Pharmacy Data

```{r}
rx |> filter(desynpuf_id == "00000B48BCF4AD29")
```

### JOIN!!!!


```{r}
beneficiary |>
  filter(desynpuf_id == "00000B48BCF4AD29") |>
  inner_join(
    rx,
    by = c(
      "desynpuf_id" = "desynpuf_id",
      "year" = "clm_year"
      )
    )
```



## Patients with/without pharmacy

- Do all patients have pharmacy? Hint: No!
  - Prove it!
- Create rx_per_beneficiary_year
- Use this to demonstrate differences in joins.
- Calculate percent male/female for each number of rx.

```{r}
rx_per_beneficiary_year <-
  rx |>
  group_by(clm_year, desynpuf_id) |>
  summarize(n = n())
rx_per_beneficiary_year
```

```{r}
rx_per_beneficiary_year |>
  group_by(clm_year) |>
  count()
```


```{r}
beneficiary |>
  mutate(
    bene_sex_ident_cd = case_when(
      bene_sex_ident_cd == 1~"male",
      bene_sex_ident_cd == 2~"female",
      TRUE~"unknown")
  ) |>
  left_join(
    rx_per_beneficiary_year,
    by = c("desynpuf_id" = "desynpuf_id", "year" = "clm_year")
  ) |>
  group_by(year, n_rx = n, bene_sex_ident_cd) |>
  count() |>
  pivot_wider(names_from = bene_sex_ident_cd, values_from = n) |>
  mutate(
    n_rx = replace_na(n_rx, 0),
    total = female+male
  ) |>
  arrange(year, n_rx)
```


# Polypharmacy

## Simple Exploration

- How many beneficiaries have 5+ prescriptions in a year?
- How many beneficiaries have 10+ prescriptions in a year?
- And why this isn't REALLY

```{r}
rx_per_beneficiary_year |>
  filter(n >= 10) |>
  group_by(clm_year) |>
  count()
```


## More Complex Exploration

- Polypharmacy is BETTER defined as a patient with 5+ prescriptions at the SAME TIME!
- Think of a better way to approach this, which we will discuss on Friday.
- You don't have to implement the code, although you may try.
- But I do want you to think about HOW to approach this.

<INSERT CODE CHUNK HERE>
