---
title: "Explore Beneficiaries"
output: html_notebook
---


```{r}
# This is the only code I'm going to give you for this.
pacman::p_load(rio, tidyverse)
beneficiary <- import("data/beneficiary_summary.rds")
```



# Beneficiaries

## Number of Beneficiaries Per Year

- Each member should have ONLY a single row per year.
- Members can appear in multiple years of data.
- Use dplyr verbs to prove/disprove the two above assertions.

```{r}
beneficiary |>
  summarize(
    n = n(),
    n_unique = length(unique(desynpuf_id))
  )
beneficiary |>
  group_by(year) |> 
  count()
```

## Min/Max Birth date

- What is the earliest date of birth in this data set?
- What is the youngest date of birth in this data set?
- You don't need to do this by year but if you do, it is fine.

```{r}
beneficiary |>
  summarize(min(bene_birth_dt))
```


```{r}
beneficiary |>
  summarize(max(bene_death_dt, na.rm = TRUE))

```


## How many beneficiaries died every year?

- Column: bene_death_dt
- Just need a count of deceased members. 
- The date doesn't matter, other than that it is not null.

```{r}
beneficiary |>
  group_by(year) |>
  summarize(n = sum(!is.na(bene_death_dt)))

beneficiary |>
  filter(!is.na(bene_death_dt)) |>
  group_by(year) |>
  summarize(n = n())

```

## Is the number of beneficiaries with diabetes increasing or decreasing?

- Count of beneficiaries per year with diabetes.
- Column: sp_diabetes
- You can look in the codebook to read the documentation, but 1 is Yes and 2 is no and this is true for all of these chronic disease T/F columns.

```{r}
beneficiary |>
  filter(sp_diabetes == 1) |>
  group_by(year) |>
  count()
```

How could you edit data/import_beneficiary_summary.R to replace 1/2 values with Yes/No values?

## Amount spent on beneficiaries

- What is the min/max/average amount paid by Medicare for inpatient events?
- As above, stratify (group_by) by year.
- Column: medreimb_ip


```{r}
beneficiary |>
  group_by(year) |> 
  summarize(
    min_paid = min(medreimb_ip),
    max_paid = max(medreimb_ip)
  )
```

```{r}
beneficiary |>
  mutate(
    sp_diabetes = case_when(sp_diabetes == 1~"Yes",
                            sp_diabetes == 2~"No",
                            TRUE~"ERROR")
  )
```

