---
title: "Self Joining"
output: html_notebook
---

```{r}
pacman::p_load(lubridate, rio, tidyverse)
beneficiary <- import("data/beneficiary_summary.rds")
## I just noticed our import script doesn't conver the discharge
## date to a true date. This mutate function fixes that & makes
## it easier to use.
inpatient <-
  import("data/inpatient_claims.rds") |>
  mutate(nch_bene_dschrg_dt = ymd(nch_bene_dschrg_dt))
```

# Self Joining

- Self joining is when a table is joined onto itself.
- The inpatient data contains an excellent example.
- Hospitals are often judged on the percentage of inpatient discharges which have a readmission within 30 days. A lower readmission rate is better. 
  - Usually the specifics are a little more detailed than that, but for demonstrative purposes, this is adequate.

And hospital readmissions is a GREAT way to demonstrate self joins.

- Each claim in inpatient has a column called `nch_bene_dschrg_dt` which is the discharge date from hospital.
- We will use the `clm_from_dt` as the admission date.
- For each discharge, I want to join inpatient on itself for EVERY admission within 30 days.
- The `select` statements below limit our result set to include ONLY those columns which are actually useful for this analysis.
  - Don't keep stuff you don't need. It will only get in the way when you self join.
  - Because you will have multiple columns with nearly identical names prepended by x and y.

```{r}
readmit <-
  ## This first use of inpatient gives us our anchor admission dates.
  inpatient |>
  select(
    desynpuf_id,
    clm_id,
    nch_bene_admssn_dt = clm_from_dt,
    nch_bene_dschrg_dt
  ) |>
  mutate(thirty_days_dschrg_dt = nch_bene_dschrg_dt + 30) |>
  ## VERY IMPORTANT TO USE LEFT JOIN!!!
  left_join(
    inpatient |>
      select(
        ## Note how I RENAME columns!!!
        desynpuf_id,
        readmit_clm_id = clm_id,
        nch_bene_readmssn_dt = clm_from_dt
      ),
    by = join_by(
      desynpuf_id,
      y$nch_bene_readmssn_dt >= x$nch_bene_dschrg_dt,
      y$nch_bene_readmssn_dt <= x$thirty_days_dschrg_dt
    )
  ) |>
  mutate(
    readmssn = case_when(
      is.na(nch_bene_readmssn_dt) ~ FALSE,
      TRUE ~ TRUE
    ),
    days_to_readmit = nch_bene_readmssn_dt - nch_bene_dschrg_dt
  )

readmit
```

- And it would seem that all is well.
- But not quite.


```{r }
dim(inpatient)
dim(readmit)
```

- Our readmit data has MORE rows than inpatient.
- Some patients had more than one readmission within 30 days of the anchor date.- We can fix this.

```{r }
readmit <-
  readmit |>
  group_by(desynpuf_id, clm_id, nch_bene_admssn_dt, nch_bene_dschrg_dt) |>
  summarize(
    n_readmits = sum(readmssn),
    readmssn = max(readmssn),
    min_days_to_readmit = min(days_to_readmit)
  )
```

Now, let's test it.

```{r }
dim(inpatient)
dim(readmit)
```

- The number of rows matches!!!
- So now we can answer our actual question!

```{r}
readmit |>
  group_by(readmssn) |>
  count() |>
  ungroup() |>
  mutate(p = 100 * n / sum(n))
```

- Over 10 percent of admissions have a readmission within 30 days.
- If we dug deeper we should find the risk of this happening goes up with age, comorbidities, etc.